---
title: "availability calc"
author: "Leslie Roberson"
date: "28/12/2021"
output: html_document
---

## Susceptibility (S)

step 1: calculate availability 

```{r setup, include=FALSE}

library(tidyverse)
library(here)
library(raster) # install.packages("raster")
library(rgdal) # install.packages("rgdal")
library(sf)
library(sp)
library(scales)

knitr::opts_chunk$set(echo = TRUE)
```

## raster math (overlay) in R

loop through all species and gears

```{r loop overlay}
## get spp rasters
dir <- here("_data/spatial/species_rasters/proj_clipped_allprobs") # 161 Mb - git ignore, stored locally
spp_rasts <- list.files(path = dir, pattern = "*.tif$", full.names = TRUE) 
spp_rasts_short <- list.files(path = dir, pattern = "*.tif$", full.names = FALSE)

##<><> get effort rast
#eff_rast <- raster::raster(here("_data/spatial/effort_rasters//GN_smth95_scaled.tif"))
#eff_rast <- raster::raster(here("_data/spatial/effort_rasters/LL_smth95_scaled.tif"))
eff_rast <- raster::raster(here("_data/spatial/effort_rasters/PS_smth95_scaled.tif"))

#<><> get the extent of one of hte species rasts
spp.extent <- raster::raster("_data/spatial/species_rasters/Fis_20033.tif")
## resample effort rast so extent exactly matches the species 
# don't put this in the loop or it'll do it 405 times (resample is slow)
eff_rast <- resample(eff_rast, spp.extent)

for(i in 1:length(spp_rasts)){
  # get rasters
  temp <- raster(spp_rasts[[i]]) 
  # get species name
  spp_name <- sub('.tif', '', spp_rasts_short[i]) 
  # multiply rasters
  avail_raster <- temp*eff_rast
  ## replace na vals with 0
  avail_raster[is.na(avail_raster[])] <- 0
  writeRaster(avail_raster, 
              #file = paste0("/Users/rob713/github/indianocean_bycatchPSA_ms/_data/spatial/availability_rasters/GN/",
              #file = paste0("/Users/rob713/github/indianocean_bycatchPSA_ms/_data/spatial/availability_rasters/LL/",
              file = paste0("_data/spatial/availability_rasters/PS/",
                           spp_name,"_avail.tiff",sep = ""), # name each file with the SpeciesID & save as geotiff
              options = c("COMPRESS=deflate","tiled=Yes"), overwrite = TRUE)
}

```

## get availability score for each species

Sum availability across all cells 
Make a df with all species
remove species with 0 availability or 0 encounterability
scale availability for each species, so availability matches other S parameters (on 1-3 continuous scale)
Scale 1-3 across all gears for availability, so they're comparable (because fishing effort is scaled together in the A calculation - same effort unit for all gears)

```{r scale 1-3}

### Make empty df to store results
result_df <- data.frame()

##<><>< First sum availability across all cells for each species

#<><><><><> Loop through all species
## REPEAT THIS FOR EACH GEAR
## get availability rasters 
dir <- here("_data/spatial/availability_rasters/LL")
#dir <- here("_data/spatial/availability_rasters/PS")
#dir <- here("_data/spatial/availability_rasters/GN") 
avail_rasts <- list.files(path = dir, pattern = "*.tif$", full.names = TRUE) 
length(avail_rasts) # 405
avail_rasts_short <- list.files(path = dir, pattern = "*.tif$", full.names = FALSE)

GEAR <- "LLT"
#GEAR <- "PST"
#GEAR <- "GND"

### Make empty df to store results
cells_df <- data.frame()
availsumm_df <- data.frame()

## loop through rasters and store all summary values in a single dataframe, with a column for gear
for(i in 1:length(avail_rasts)){
  # get rasters
  temp <- raster::raster(avail_rasts[i])
  # get species name
  species_id <- gsub("\\_avail.tif", "", avail_rasts_short[i])
  # convert raster to df
  temp_df <- raster::as.data.frame(temp, xy = TRUE, na.rm=TRUE) 
  # rename availability column
  names(temp_df)[3] <- "A"
  # add species id and gear cols
  temp_df$species_id <- species_id
  temp_df$GEAR <- GEAR
  # remove cells that don't overlap with fishing
  temp_df <- temp_df %>% filter(A>0)
  # add the full cell based availability to a results df
  cells_df = rbind(cells_df, temp_df)
  # count cells that overlap with fishing
  ncells_overlap <- temp_df %>% summarise(ncells_overlap=nrow(temp_df))
  ## sum all A values
  sumA <- temp_df %>% summarise(sumA=sum(A))
  ## add to df
  outputsumm <- data.frame(species_id, sumA, ncells_overlap, GEAR)
  availsumm_df = rbind(availsumm_df, outputsumm)
}

## save cell dfs separte by gear (for mapping)
glimpse(cells_df)
summary(cells_df) # No NAs or A==0

gnd.cells <- cells_df
fwrite(gnd.cells, here("_results/Susceptibility/Availability/GND_availcells.csv"))
pst.cells <- cells_df
fwrite(pst.cells, here("_results/Susceptibility/Availability/PST_availcells.csv"))
llt.cells <- cells_df
fwrite(llt.cells, here("_results/Susceptibility/Availability/LLT_availcells.csv"))
## add these to gitignore - too big for github

## combined and scale A for summary df
glimpse(availsumm_df)
gnd.summ <- availsumm_df
pst.summ <- availsumm_df
llt.summ <- availsumm_df

allgears.summ <- bind_rows(gnd.summ, pst.summ, llt.summ)
glimpse(allgears.summ)

## scale A (across all gears)
allgears.summ  <- allgears.summ  %>%
  mutate(A_scaled = scales::rescale(sumA, to= c(1,3)))

## checks
summary(allgears.summ)
unique(allgears.summ$GEAR)

write_csv(allgears.summ, here("_results/Susceptibility/Availability/sum_availcells_allgears.csv"))

```

