---
title: "conservation status and number of fisheries species is susceptible to"
author: "Leslie Roberson"
date: "08/02/2022"
output: html_document
---

Make plot showing Vulnerability, IUCN status and number of fisheries species is susceptible to

```{r setup, include=FALSE}

library(tidyverse)
library(janitor)
library(pillar)
library(readxl)
library(ggpubr)
library(cowplot)
library(grid)
library(gridExtra)
library(ggridges)

knitr::opts_chunk$set(echo = TRUE)
```

# Make plot dat

Get IUCN statuses from iucnredlist.org (version 2021.3)
Multiple match process on species names

NB the following species ended in a manual search (where names from AquaMaps database don't match Red List name):
 1 Chiloscyllium hasseltii     EN
 2 Dentiraja confusus         CR 
 3 Mobula mobula      EN         
 4 Pseudobatus percellens    EN  
 5 Himantura granulata     VU     
 6 Stegostoma fasciatum EN        
 7 Delphinus delphis tropicalis   LC  ** taxonomy debated
 8 Balaenoptera brydei      DD ** taxonomy debated on this one too   
 9 Natator depressa       DD         
10 Centroscymnus crepidater  NT  

```{r make plot dat}

##<>>< Make IUCN status df

# Don't need all_other_fields.csv (extra conservation info)
# Don't need the taxonomy file either
assess_raw <- read_csv("_data/IUCN_redlist_species_data_2021_3/assessments.csv")
glimpse(assess_raw)
assess <- assess_raw %>% clean_names(case = "snake") %>% dplyr::select(c(internal_taxon_id, scientific_name, redlist_category))

## load vulnerability summary 
v.scores <- read_csv(here("_results/Vulnerability/V_scores.csv"))
glimpse(v.scores)
## fix any capitalization issues in AM database
v.scores <- v.scores %>%
  mutate(scientific_name=stringr::str_to_sentence(scientific_name))

## match on scientific name
V.dat <- left_join(v.scores, assess, by="scientific_name")
glimpse(V.dat)
V.dat %>% filter(is.na(redlist_category)) %>% distinct(scientific_name) # 10 unmatched

## Manual search
V.dat <- V.dat %>%
  mutate(redlist_category = factor(redlist_category, 
                                   levels=c("Critically Endangered", "Endangered", "Vulnerable", "Near Threatened", "Least Concern", "Data Deficient")),
         scientific_name=stringr::str_replace(scientific_name, "Delphinus delphis tropicalis", "Delphinus delphis"),
         redlist_category=case_when(
           scientific_name=="Dentiraja confusus" ~ "Critically Endangered",
           scientific_name %in% c("Chiloscyllium hasseltii", "Mobula mobula", "Pseudobatus percellens", "Stegostoma fasciatum") ~ "Endangered",
           scientific_name =="Himantura granulata" ~ "Vulnerable",
           scientific_name=="Centroscymnus crepidater" ~ "Near Threatened",
           scientific_name=="Delphinus delphis" ~ "Least Concern",
           scientific_name %in% c("Natator depressa", "Balaenoptera brydei") ~ "Data Deficient",
           TRUE~as.character(redlist_category)),
         iucn_sciname=case_when(
           scientific_name=="Chiloscyllium hasseltii"~"Chiloscyllium hasselti",
           scientific_name=="Dentiraja confusus"~"Dentiraja confusa",
           scientific_name=="Mobula mobula"~"Mobula mobular",
           scientific_name=="Pseudobatus percellens"~"Pseudobatos percellens",
           scientific_name=="Himantura granulata"~"Urogymnus granulatus",
           scientific_name=="Stegostoma fasciatum"~"Stegostoma tigrinum",
           scientific_name=="Natator depressa"~"Natator depressus",
           scientific_name=="Centroscymnus crepidater"~"Centroselachus crepidater",
           scientific_name=="Balaenoptera brydei"~"Balaenoptera omurai",
           TRUE~as.character(scientific_name)
         ))
glimpse(V.dat)
V.dat %>% filter(is.na(redlist_category)) # none

V.dat <- V.dat %>%
  group_by(species_id) %>%
  mutate(n_gears=n_distinct(gear)) %>%
  ungroup()

V.dat %>% distinct(species_id, n_gears) %>% tabyl(n_gears)
# n_gears   n   percent
#       1  64 0.2006270
#       2  99 0.3103448
#       3 156 0.4890282

write_csv(V.dat, "_results/figures/fig5_iucn_dat.csv")

```

## Plot 2 panels

IUCN status on y axis and V score on X axis 
number of gears susceptible to
color corresponds to ngears

```{r scatter}

dat <- read_csv(here("_results/figures/fig5_iucn_dat.csv"))
glimpse(dat)
pdat <- dat %>% 
  group_by(species_id, n_gears, redlist_category) %>%
  summarise(avgV=mean(V), # mean V score across the 3 gears
            avgV123=mean(V123))
summary(pdat$avgV123)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 1.822   2.557   2.836   2.849   3.153   3.702

##<><>> prep data

pdat$n_gears <- as_factor(pdat$n_gears)
# color scheme that's different frmo the other plots
pal_ngears <- c("cadetblue1", "dodgerblue1", "royalblue4")
pdat$redlist_category <- as_factor(pdat$redlist_category)
pdat$redlist_category <- fct_relevel(pdat$redlist_category, levels=c("Critically Endangered", "Endangered", "Vulnerable", "Near Threatened", "Least Concern", "pdata Deficient"))

## plot
p.density <- pdat %>%
  ggplot() +
  geom_density_ridges(aes(x=avgV, y=redlist_category, fill=n_gears, color=n_gears), alpha=0.75) +
  labs(y="", x="Mean Vulnerability score", fill="", color="") +
  scale_y_discrete(limits=rev) +
  scale_x_continuous(limits=c(1.5,4.5), breaks=seq(2,4,0.5)) +
  scale_color_manual(values=pal_ngears) + 
  scale_fill_manual(values=pal_ngears) + 
  theme_bw() +
  theme(legend.position="top",
        panel.grid = element_blank()) +
  guides(color="none")
p.density

##<><>< Add a bar plot to the right side
p.bar <- pdat %>%
  group_by(redlist_category, n_gears) %>%
  summarise(n_spp=n_distinct(species_id)) %>%
  ggplot(aes(fill=n_gears, x=redlist_category, y=n_spp)) +
  geom_bar(width=0.4, position="stack", stat="identity") +
  scale_color_manual(values=pal_ngears) + 
  scale_fill_manual(values=pal_ngears) + 
  scale_x_discrete(limits=rev) +
  scale_y_continuous(breaks=seq(0,125,25)) +
  labs(x="", y="Number of species", fill="Number of gears") +
  coord_flip() +
  theme_bw() +
  theme(legend.position="top",
        panel.grid = element_blank(),
        axis.ticks.y=element_blank()) +
  guides(color="none")
p.bar

##<>><>< make common legend
legend_gears <- get_legend(
  p.bar + 
    guides(fill = guide_legend(nrow = 1)) +
    theme(legend.position = "top")
)
legend_gears

#ggarrange(legend_gears, p.density + guides(fill="none"), p.bar + rremove("y.text") + guides(fill="none"),
#         labels=c("A", "B"),
#         nrow=2)

iucn.grid <- grid.arrange(legend_gears,
             p.density + theme(legend.position="none"),
             p.bar + theme(legend.position="none",
                                       axis.title.y=element_blank(),
                                       axis.text.y=element_blank(),
                                       axis.ticks.y=element_blank()),
            ncol=2, nrow=2,
            layout_matrix = rbind(c(1,1), c(2,3)),
          widths = c(3.5, 2), heights = c(0.2, 2.5))
iucn.grid

ggsave(iucn.grid,
       width=6, height=5, units="in",
       dpi=600,
       filename="_results/figures/fig5_iucngrid.eps",
       device=cairo_ps)

```

Explore plot variations

```{r plot explore}
##<>< plot explore
ptest <- dat %>%
  ggplot() +
  geom_density(aes(y=V123, fill=n_gears, color=n_gears), alpha=0.8) +
  #geom_jitter(aes(x=redlist_category, y=V123, color=n_gears), alpha=0.8) +
              #colour="black",pch=21) + # dots with outline
  #stat_chull(aes(x=redlist_category, y=V123, color = n_gears, fill = n_gears), alpha = 0.1, geom = "polygon") + # this would also be cool for just one scatterplot with lots of dots, to show groups
  labs(x="", y="Vulnerability score", fill="Number of gears", color="") +
  scale_color_manual(values=pal_ngears) + 
  scale_fill_manual(values=pal_ngears) + 
  #scale_fill_grey(start=0.8, end=0.1) +
  #scale_color_manual(values=pal_taxgrp) +
  #scale_size_area(max_size=4) + # size of dots makes it too cluttered
  #scale_x_discrete(limits=rev) + # so CR is on top
  coord_flip() +
  facet_wrap(~redlist_category, nrow=2) + # how do I make the density curves on 3 different levels so they're not overlapping?
  #facet_grid(n_gears~redlist_category) + # nah too many boxes
  theme_bw() +
  theme(legend.position="top",
        panel.grid = element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.background =element_rect(fill="white"))
ptest ## hmmm no it's just a mess of huge dots 

ggscatterhist(
  dat, x="redlist_category", y="V123", color="n_gears", size = 2, alpha = 0.6,
  palette = c("#00AFBB", "#E7B800", "#FC4E07"),
  margin.plot = "boxplot",
  ggtheme = theme_bw()
  ) # that's cool but it only really works for a normal continuous by continuous scatter

```

## Plot V1

grouped stacked barplot with the IUCN grups and number of species in 1, 2, or 3 fisheries

```{r bar plot}

glimpse(V.dat)
V.dat$n_gears <- as_factor(V.dat$n_gears)
V.dat$redlist_category <- as_factor(V.dat$redlist_category)
V.dat$redlist_category <- fct_relevel(V.dat$redlist_category, levels=c("Critically Endangered", "Endangered", "Vulnerable", "Near Threatened", "Least Concern", "Data Deficient"))

iucn.plot <- V.dat %>%
  distinct(species_id, n_gears, redlist_category) %>%
  ggplot(aes(fill=n_gears, x=redlist_category)) +
  geom_bar(width=0.75, position=position_dodge(0.75), stat="count", color="black", size=0.4) +
  labs(x="", y="Number of species", fill = "Number of\nfishing gears") +
  scale_fill_grey(start=0.8, end=0.1) +
  scale_x_discrete(limits=rev) +
  scale_y_continuous(limits=c(0,45), breaks=seq(0,40,10)) +
  coord_flip() +
  theme_bw() +
  theme(panel.grid = element_blank(),
        legend.text=element_text(size=6)) +
  guides(fill=guide_legend(reverse=TRUE))
iucn.plot

pdf("_results/figures/fig_IUCN/10feb_grpedbarplot_grey.pdf", width=5, height=2.5, bg="white")
print(iucn.plot)
dev.off()

```

